### 2017-11-09 paper/sections/2_parsplice-keyspace-analysis.tex
paper: addresses #9
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

%\section{ParSplice Keyspace Analysis}
%
%We demonstrate these contributions by changing the input to our molecular
%dynamics simulation, which changes the keyspace access patterns. 
%
%\subsection{Structured Access Regimes}
%
%Figures~\ref{fig:keyspace-analysis_size},~\ref{fig:keyspace-analysis_locality} and~\ref{fig:keyspace-analysis_throughput}.
%%shows ``access regimes", or groups of accesses
%%to the same subset of keys:
%\begin{itemize}
%  \item change immediately, different sizes
%  \item monotonically increasing
%  \item random access to a single key
%  \item early keys accessed more
%\end{itemize}
%
%\noindent\textbf{Conclusion}: unique patterns of a real HPC application
%
%\subsection{Cache Size Analysis}
%
%Figure~\ref{fig:keyspace-analysis_cachesize}.
%
%\noindent\textbf{Conclusion}: workload phases (high request rate to a small
%number of keys and then low request rate to a large number keys) need a dynamic
%load balancing policy.
%
%\subsection{Overfitting Policies}
%
%Figures~\ref{fig:keyspace-analysis_cachesize-dynamic} and~\ref{fig:keyspace-analysis_cachesize-caveats}.
%
%\noindent\textbf{Side Idea}: we need to figure out what ParSplice memory is
%sensitive to: max usage or usage over time.\\
%
%\noindent\textbf{Conclusion}: dynamic policies absorb the cost of a high read
%request rate for a 2.5 hour run, but it is infeasible to do this for every
%combination of system setup, job lengths, parsplice parameters.
%
%\subsubsection*{Discussion}
%Why don't we just an LRU cache:
%\begin{itemize}
%  \item we can do better (workload is structured and has locality)
%  \item finding the size of the cache is hard
%  \item hotspots dissipate too quickly
%\end{itemize}

@@

We use ParSplice to simulate the evolution of metallic nanoparticles that grow
from the vapor phase. As the run progresses, the energy landscape of the system
becomes more complex. Two domain factors control the number of entries in the
data store: the growth rate and the temperature. The growth rate controls how
quickly new atoms are added to the nanoparticle: fast growth rates lead to
non-equilibrium conditions, and hence increase the number of states that can be
visited. However, as the particle grows, the simulation slows down because the
calculations become more expensive, limiting the rate at which new states are
visited. On the other hand, the temperature controls how easily a trajectory
can jump from state to state; higher temperatures lead to more frequent
transitions.

The nanoparticle simulation stresses the data store architecture of ParSplice.
It visits more states than other input decks because the system uses a cheap
potential, has a small number of atoms, and operates in a complex energy
landscape with many accessible states. Changing growth rates and temperature
alters the size, shape, and locality of the data store keyspace. Lower
temperatures and smaller growth rates create hotter keys with smaller keyspaces
as many segments are generated in the same set of states before the trajectory
can escape to a new region of state space.

@@

\subsubsection*{Entropy increases over time} The reads per second in
Figure~\ref{fig:motivation-regimes} show that the number of requests decreases
and the number of active keys increases over time. The resulting key access
imbalance for the two growth rates in Figure~\ref{fig:motivation-regimes} are
shown in Figure~\ref{fig:methodology-keys}, where reads are plotted for each
unique state, or key, along the \(x\) axis. Keys are more popular than others
(up to \(5\times\)) because worker tasks start generating states with different
coordinates later in the run (this motivates
Section~\S\ref{sec:cache-management-using-domain-specific-knowledge}).  The
growth rate, temperature, and number of workers have a predictable effect on
the structure of the keyspace.  Figure~\ref{fig:methodology-keys} shows that
the number of reads changes with different initial conditions (\(Delta\)), but
that the spatial locality of key accesses is similar ({\it e.g.}, some keys are
still \(5\times\) more popular than others).
Figure~\ref{fig:motivation-regimes} shows how entropy for different growth
rates has temporal locality, as the reads per second for \(\Delta_2\) looks
like the reads per second for \(\Delta_1\) stretched out along the time axis.
Trends also exist for temperature and number of workers but are omitted here
for space. This structure means that we can learn the regimes and adapt the
storage system (this motivates
Section~\S\ref{sec:cache-management-using-domain-specific-knowledge}).

>>>>

%
%We demonstrate these contributions by changing the input to our molecular
%dynamics simulation, which changes the keyspace access patterns. 
%
%\subsection{Structured Access Regimes}
%
%Figures~\ref{fig:keyspace-analysis_size},~\ref{fig:keyspace-analysis_locality} and~\ref{fig:keyspace-analysis_throughput}.
%%shows ``access regimes", or groups of accesses
%%to the same subset of keys:
%\begin{itemize}
%  \item change immediately, different sizes
%  \item monotonically increasing
%  \item random access to a single key
%  \item early keys accessed more
%\end{itemize}
%
%\noindent\textbf{Conclusion}: unique patterns of a real HPC application
%
%\subsection{Cache Size Analysis}
%
%Figure~\ref{fig:keyspace-analysis_cachesize}.
%
%\noindent\textbf{Conclusion}: workload phases (high request rate to a small
%number of keys and then low request rate to a large number keys) need a dynamic
%load balancing policy.
%
%\subsection{Overfitting Policies}
%
%Figures~\ref{fig:keyspace-analysis_cachesize-dynamic} and~\ref{fig:keyspace-analysis_cachesize-caveats}.
%
%\noindent\textbf{Side Idea}: we need to figure out what ParSplice memory is
%sensitive to: max usage or usage over time.\\
%
%\noindent\textbf{Conclusion}: dynamic policies absorb the cost of a high read
%request rate for a 2.5 hour run, but it is infeasible to do this for every
%combination of system setup, job lengths, parsplice parameters.
%
%\subsubsection*{Discussion}
%Why don't we just an LRU cache:
%\begin{itemize}
%  \item we can do better (workload is structured and has locality)
%  \item finding the size of the cache is hard
%  \item hotspots dissipate too quickly
%\end{itemize}

@@

We use ParSplice to simulate the evolution of metallic nanoparticles that grow
from the vapor phase.  This simulation stresses the storage hierarchy more than
other input decks because it uses a cheap potential, has a small number of
atoms, and operates in a complex energy landscape with many accessible states.
As the run progresses, the energy landscape of the system becomes more complex
and more states are visited.  Two domain factors control the number of entries
in the data store: the growth rate and the temperature. The growth rate
controls how quickly new atoms are added to the nanoparticle: fast growth rates
lead to non-equilibrium conditions, and hence increase the number of states
that can be visited.  However, as the particle grows, the simulation slows down
because the calculations become more expensive, limiting the rate at which new
states are visited.  On the other hand, the temperature controls how easily a
trajectory can jump from state to state; higher temperatures lead to more
frequent transitions but temperatures that are too high result in meaningless
simulations because trajectories have so much energy that they are equally
likely to visit any random state.

Changing growth rates and temperature alters the size, shape, and locality of
the data store keyspace. Lower temperatures and smaller growth rates create
hotter keys with smaller keyspaces as many segments are generated in the same
set of states before the trajectory can escape to a new region of state space.
In Figure~\ref{fig:motivation-regimes}, the \(\Delta_1\) growth rate adds atoms
every 100K microseconds while the \(\Delta_2\) growth rate adds atoms every 1
million microseconds. So \(\Delta_2\) has a smaller growth rate resulting in
hotter keys (line on \(y1\) axis) and a smaller keyspace (dots on the \(y2\)
axis).  Values smaller than \(\Delta_2\)'s growth rate or a temperature of 400
degrees results in very little database activity because state transitions take
too long. Similarly, values larger than \(\Delta_1\)'s growth rate or a
temperature of 4000 degrees result in an equally meaningless simulation as
transitions are unrealistic.

@@

\subsubsection*{Entropy increases over time} The reads per second in
Figure~\ref{fig:motivation-regimes} show that the number of requests decreases
and the number of active keys increases over time. The resulting key access
imbalance for the two growth rates in Figure~\ref{fig:motivation-regimes} are
shown in Figure~\ref{fig:methodology-keys}, where reads are plotted for each
unique state, or key, along the \(x\) axis. Keys are more popular than others
(up to \(5\times\)) because worker tasks start generating states with different
coordinates later in the run (this motivates
Section~\S\ref{sec:cache-management-using-domain-specific-knowledge}).  The
growth rate, temperature, and number of workers have a predictable effect on
the structure of the keyspace.  Figure~\ref{fig:methodology-keys} shows that
the number of reads changes with different initial conditions (\(\Delta\)), but
that the spatial locality of key accesses is similar ({\it e.g.}, some keys are
still \(5\times\) more popular than others).
Figure~\ref{fig:motivation-regimes} shows how entropy for different growth
rates has temporal locality, as the reads per second for \(\Delta_2\) looks
like the reads per second for \(\Delta_1\) stretched out along the time axis.
Trends also exist for temperature and number of workers but are omitted here
for space. This structure means that we can learn the regimes and adapt the
storage system (this motivates
Section~\S\ref{sec:cache-management-using-domain-specific-knowledge}).
